Field-informed Reinforcement Learning for Learning Large-Scale Collective Tasks

### 2023-05-17 paper.tex

Finally, to maximise the coverage, 
  we define a reward function that favours the maximum distance between the agents
  equal to the coverage range $\omega$. This will minimise multiple agents covering a common area.
%
This means that the average distance will tend towards the one expressed by the viewing range of each node:
% 
\begin{equation*}
R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
\end{equation*}
where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
%
%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
%
\subsection{Training Phase}
The learning process was divided into 100 episodes, each consisting of 200 steps, resulting in a total of 20,000 experiences.
%
For each episode, 
 the 25 agents are semi-randomly positioned on a grid without knowing the correct position of the phenomenon, 
 but that is fixed in the top right corner. 
The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
%
In this case, we chose to use an exponential epsilon decay, defined as:
\begin{equation*}
\epsilon = \epsilon_{min} + (\epsilon_{max} - \epsilon_{min}) \cdot e^{-\lambda \cdot e}
\end{equation*}
where e is the current episode number. 
This leads to a high number of random actions at the beginning and gradually shifts towards exploitation in the later episodes. 
%
In our training process, 
 we set $\epsilon_{min} = 0.02$, $\epsilon_{max} = 0.99$, and $\lambda = 0.1$.
$\gamma$ was set to 0.99, 
 as we want to give more value to future returns, 
 aiming to achieve good coverage by continuously tracking the phenomenon. 
The neural network structure used consists of a layer of SuperGAT~\cite{DBLP:journals/corr/abs-2204-04879} -- A \ac{GNN} based on attention mechanisms -- and a layer of MLP. 
The hidden size was set to 256.
%
As the reward function is defined as a regret, 
 we decided to use the Huber loss function, 
 which is a combination of the $L_1$ and $L_2$ loss functions.
%
The loss function is defined as:
\begin{equation*}
L = \begin{cases}
  \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise.} 
 \end{cases}
\end{equation*}
This function is used to penalise the agent if the action taken is too far from the optimal action.
%
We use the RMSprop optimiser with a learning rate of 0.0001.
%
Finally, we use a replay buffer of size 1000 to store the graph experiences and a batch size of 32.
\subsection{Test phase}
For the evaluation, 
 we explore three different types of scenarios.
%
The first type considers only a single, stationary phenomenon like the one used in the training phase. 
%
The second type considers multiple phenomena, requiring the agents to make a decision about which phenomena to cover. 
%
Finally, the third type considers a single moving phenomenon which the agents have to follow throughout the simulation. 
%
In all three types of scenarios, the shape and underlying distribution are initially unknown to the agents but remain constant for the entire duration of the simulation. 
%
We generated random scenarios for each type of scenario. 
%
Examples of such types are shown in~\Cref{fig:scenarios}. %\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}

>>>>

Finally, to maximise the coverage, 
  we define a reward function that favours the maximum distance between the agents
  equal to the coverage range $\omega$. This will minimise multiple agents covering a common area.
This means that the average distance will tend towards the one expressed by the viewing range of each node:
\begin{equation*}
R^{C}_{v} = 1 - \frac{d_{min}}{\omega}
\end{equation*}
where $d_{min}$ describes and minimum distance of the agent to its neighbourhood. 
%\lukas{should this be $1-R^C_v$? or do we want to favour the highest average distance?}
The final reward function for an agent $v$ is defined as:
\begin{equation*}
R_{v} = \frac{R^a_{v} + R^{\mathcal{N}}_{v} + R^{C}_{v}}{3}
\end{equation*}
Specifically, we decided to express the signal as a \emph{regret}
  as it is a more general measure of the quality of the action taken by the agent.
It is simplify evaluated as:
\begin{equation*}
R^e_{v} = R_v - 1
\end{equation*}
\subsection{Training Phase}
The learning process was divided into 100 episodes, each consisting of 200 steps, resulting in a total of 20,000 experiences.
For each episode, 
 the 25 agents are semi-randomly \lukas{what is semi-random? are the position fixed but the agents random? or is it in a specific layout (e.g. lattice) but with some randomness of the actual position?} positioned on a grid without knowing the correct position of the phenomenon, 
 but that is fixed in the top right corner.  \lukas{so for all 20,000 experiences, the phenomena is exactly as shown in fig 2a during training?}
The feature set used by the \ac{GNN} created for each node consists of the vector computed by the aggregated program and the value of the local sensor $\zeta$.
In this case, we chose to use an exponential epsilon decay, defined as:
\begin{equation*}
\epsilon = \epsilon_{min} + (\epsilon_{max} - \epsilon_{min}) \cdot e^{-\lambda \cdot e}
\end{equation*}
where e is the current episode number. 
This leads to a high number of random actions at the beginning and gradually shifts towards exploitation in the later episodes. 
In our training process, 
 we set $\epsilon_{min} = 0.02$, $\epsilon_{max} = 0.99$, and $\lambda = 0.1$.
$\gamma$ was set to 0.99, 
 as we want to give more value to future returns, 
 aiming to achieve good coverage by continuously tracking the phenomenon. 
The neural network structure used consists of a layer of SuperGAT~\cite{DBLP:journals/corr/abs-2204-04879} -- A \ac{GNN} based on attention mechanisms -- and a layer of MLP. 
The hidden size was set to 256.
As the reward function is defined as a regret, 
 we decided to use the Huber loss function~\cite{??}, 
 which is a combination of the $L_1$ and $L_2$ loss functions.
The loss function is defined as:
\begin{equation*}
L = \begin{cases}
  \frac{1}{2} (y - \hat{y})^2 & \text{if } |y - \hat{y}| < \delta \\
  \delta (|y - \hat{y}| - \frac{1}{2} \delta) & \text{otherwise.} 
 \end{cases}
\end{equation*}
This function is used to penalise the agent if the action taken is too far from the optimal action. \lukas{whats $y, \hat{y}$ and $\delta$?}
We use the RMSprop optimiser with a learning rate of 0.0001.
Finally, we use a replay buffer of size 1000 to store the graph experiences and a batch size of 32.
\subsection{Test phase}
For the evaluation, 
 we explore three different types of scenarios.
The first type considers only a single, stationary phenomenon like the one used in the training phase. 
The second type considers multiple phenomena, requiring the agents to make a decision about which phenomena to cover. 
Finally, the third type considers a single moving phenomenon which the agents have to follow throughout the simulation. 
In all three types of scenarios, the shape and underlying distribution are initially unknown to the agents but remain constant for the entire duration of the simulation. 
We generated random scenarios for each type of scenario. 
Examples of such types are shown in~\Cref{fig:scenarios}. %\lukas{can we create a figure with at least 3 sample scenarios (1 stationary, 2 split, 1 moving (with a line indicating its trajectory))?}
